---
title: "Stand_Summary_Build"
output: html_notebook
---

Building up stand summary tool functionality before moving to shiny app.  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
knitr::opts_knit$set(root.dir = "..")

library(tidyverse)
library(sf)
library(leaflet)
library(htmltools)

```


Which data do we want to display and summarize? tree density, BA, CC, regen, surface fuels, shrub cvr, treatment history (treated, untreated), density by size class.  

```{r}
## general data with treatment indicator
gen <- read.csv("data/Stand/GeneralPlot.csv") %>%
  dplyr::select(plot = Plot.Name, forest = Forest, loc = Location, treated = Treated) %>%
  filter(!is.na(plot)) %>%
  arrange(plot)
## cover data
cover <- read.csv("data/Stand/CoverData.csv") %>%
  dplyr::select(plot = PlotName, live_cc = LiveOverstory, dead_cc = DeadOverstory, 
                shrub_c = ShrubCover, litter = Litter, woody_debris = WoodyDebris) %>%
  filter(!is.na(plot)) %>%
  arrange(plot)
## Tree data
trees <- read.csv("data/Stand/TreeData.csv") %>%
  dplyr::select(plot = PlotName, tree = TreeNumber, species = Species, status = Status, 
                dbh = DBH, live_crown = LiveCrownPercent, 
                insect = InsectEvidence, insect_lev = InsectLevel) %>%
  ## idicator of live or not (including infested trees that have a green crown)
  mutate(live = ifelse(status %in% c("L", "I"), 1, 0),
         ## Calculate BA
         ba = pi * (dbh/2)^2 / 10000, # returns square meters
         ## Set some size classes
         class = case_when(dbh < 20 ~ 1,
                 dbh >= 20 & dbh < 40 ~ 2,
                 dbh >= 40 & dbh < 60 ~ 3,
                 dbh >= 60 & dbh < 80 ~ 4,
                 dbh >= 80 & dbh < 100 ~ 5,
                 dbh >= 100 & dbh < 120 ~ 6,
                 dbh >= 120 ~ 7)) 

## Summarise tree data by plot for dead and alive
## First calculate plot area as a proportion of a hectare
plot_area <- pi*12.6^2 / 10000 #12.6m radius; returns hectares

liveclass <- group_by(trees, plot) %>%
  filter(live == 1) %>%
  select(plot, class) %>%
  count(class) %>%
  mutate( tpha = round(n / plot_area, 0),
          class = paste0("class_live",class)) %>%
  select(-n) %>%
  spread(class, tpha, fill = 0) %>%
  ungroup() %>%
  mutate(tpha_live = rowSums(.[2:8]))

liveplot <- group_by(trees, plot) %>%
  filter(live == 1) %>%
  add_count(class) %>%
  summarise(ba_live = sum(ba),
            dbh_mn_live = round(mean(dbh), 0),
            dbh_max_live = round(max(dbh), 0)) %>%
  ## Put ba on per hectare basis
  mutate(ba_live = round(ba_live / plot_area, 2)) %>%
  merge(liveclass, all.x = T)

deadclass <- group_by(trees, plot) %>%
  filter(live == 0) %>%
  select(plot, class) %>%
  count(class) %>%
  mutate( tpha = round(n / plot_area, 0),
          class = paste0("class_dead",class)) %>%
  select(-n) %>% #needed or spread duplicates rows
  spread(class, tpha, fill = 0) %>%
  select(-class_deadNA) %>%
  ungroup() %>%
  mutate(tpha_dead = rowSums(.[2:8]))

deadplot <- group_by(trees, plot) %>%
  filter(live == 0) %>%
  add_count(class) %>%
  summarise(ba_dead = sum(ba),
            dbh_mn_dead = round(mean(dbh), 0),
            dbh_max_dead = round(max(dbh), 0)) %>%
  ## Put ba on per hectare basis
  mutate(ba_dead = round(ba_dead / plot_area, 2)) %>%
  merge(deadclass, all.x = T)

plot_sum <- merge(liveplot, deadplot, all = T) %>%
  filter(!is.na(plot)) %>%
  replace(., is.na(.), 0) %>%
  arrange(plot)
```

Bring in spatial data.  

```{r}
## Stand location data
d <- read.csv("data/Stand/Plot_Locations2.csv") %>%
  ## just those completed
  # filter(Completed. == "Y") %>%
  dplyr::select(plot = Plots2017, forest = Forest, loc = Location, 
         lat = Latitude, long = Longitude, utm_n = UTM_N_Corr, utm_e = UTM_E_Corr,
         elev = Elevation1) %>%
  mutate(plot = as.factor(plot)) %>%
  ## A lot of points don't appear to have been surveyed
  filter(plot %in% gen$plot) %>%
  arrange(plot)

## Convert to spatial
pts <- st_as_sf(d, coords = c("long", "lat"), crs = "+proj=longlat +datum=WGS84 +no_defs")

## AOI perimeters
district <- st_read("app_data", "SN_districts") %>%
  dplyr::select(aoi)

## Just those with points including yosemite
yose <- st_read("C:/Users/zlsteel/Documents/geodata/NPSBoundary", "nps_boundary") %>%
  filter(UNIT_NAME == "Yosemite") %>%
  mutate(aoi = "Yosemite") %>%
  dplyr::select(aoi) %>%
  st_transform(crs = st_crs(district))

keep <- st_intersects(district, pts) %>%
  apply(1, any)

aois <- district[keep,] %>%
  rbind(yose) 

## Add aoi identifier to data
d <- st_intersection(pts, district) %>%
  as.data.frame() %>%
  select(plot, forest, aoi) %>%
  merge(d, all = T) %>%
  mutate(aoi = ifelse(is.na(aoi), as.character(forest), 
                      as.character(aoi)))

```

```{r}
## Add everything to make a massive datafile
## There seems to be some points missing
# missing <- d$plot[!(d$plot %in% cover$plot)]
# missing2 <- gen$plot[!(gen$plot %in% d$plot)]
# 
# ## Which plots have locations but no data and visa versa
# plot_d <- select(gen, plot) %>%
#   mutate(data = 1)
# plot_l <- select(d, plot) %>%
#   mutate(loc = 1)
# missing3 <- merge(plot_d, plot_l, all = T) 
# write.csv(missing3, "data/Stand/missing.csv", row.names = F)

## Currently a lot of miss-matched plot names, but will fix later
d2 <- merge(d, gen[,c("plot", "treated")]) %>%
  merge(cover, all.x = T) %>%
  merge(plot_sum) %>%
  mutate(treated = ifelse(treated == "Y", "Yes", "No"))



```


Incorporate into a leaflet map.  

```{r}
leaflet(d2) %>%
  addProviderTiles(provider = "Esri.WorldImagery", group = "Aerial Imagery") %>%
  addMarkers(~long, ~lat,
             popup = paste0("PlotID: ", d2$plot, "<br>",
                            "Forest: ", d2$forest, "<br>",
                            "Location: ", d2$loc, "<br>",
                            "Treated: ", d2$treated, "<br><br>",
                            
                            "Canopy Cover (Live): ", d2$live_cc, " (%)<br>",
                            "Shrub Cover: ", d2$shrub_c, " (%)<br>",
                            "Litter: ", d2$litter, " (%)<br>",
                            "Woody Debris: ", d2$woody_debris, " (%)<br><br>",
                            
                            "Density (Live): ", d2$tpha_live, " (trees/ha)<br>",
                            "Basal Area (Live): ", d2$ba_live, " (sq m)<br>",
                            "Mean DBH (Live): ", d2$dbh_mn_live, " (cm)<br>",
                            "Max DBH (Live): ", d2$dbh_max_live, " (cm)<br><br>",
                            
                            "Density (Dead): ", d2$tpha_dead, " (trees/ha)<br>",
                            "Basal Area (Dead): ", d2$ba_dead, " (sq m)<br>",
                            "Mean DBH (Dead): ", d2$dbh_mn_dead, " (cm)<br>",
                            "Max DBH (Dead): ", d2$dbh_max_dead, " (cm)")) %>%
  addPolygons(data = aois,
              label = aois$aoi,
              highlightOptions = highlightOptions(color = "white", weight = 2,
                                                  bringToFront = TRUE))
```


What sort of figures do we want the user to be able to play with? How about boxplots with points overlayed, split by live and dead, and maybe treated and untreated.  

```{r}
library(plotly)
stand <- read.csv("data/Stand/stand_prepped.csv")

metrics <- data.frame(label = c("Canopy Cover", "Trees Density", "Basal Area", "Mean DBH", "Max DBH"), 
                      metric = c("live_cc", "tpha_live", "ba_live", "dbh_mn_live", "dbh_max_live", 
                              "dead_cc", "tpha_dead", "ba_dead", "dbh_mn_dead", "dbh_max_dead"),
                      unit = c("%", "trees/ha", "sq m", "cm", "cm", 
                               "%", "trees/ha", "sq m", "cm", "cm"),
                      status = c(rep("live",5),rep("dead",5)))

ulab <- "Canopy Cover"
plab <- paste0(ulab, " (", metrics[metrics$label == ulab, "unit"][1], ")")

vars <- filter(metrics, label == ulab) %>%
  pull(metric) %>%
  as.character()

stand2 <- select(stand, plot, aoi, treated, {{vars}}) %>%
  gather(key = metric, value = value, -plot, -aoi, -treated) %>%
  merge(metrics, by = "metric") %>%
  mutate(treat = ifelse(treated == "Yes", "Treated", "Untreated"),
         text = paste("value:",value))

p <- ggplot(stand2, aes(x = status, y = value)) +
  geom_boxplot() +
  geom_jitter() +
  facet_grid(~ treat) +
  ylab(plab) + xlab("Drought Mortality")

ggplotly(p)
```